---
title: "260677676_Assignment_4 -_MATH_525"
author: "Dan Yunheum Seol"
date: '2019 4 20 '
output:
  pdf_document: default
  html_document: default
---
# 6.6
The  file  azcounties.dat  gives  data  from  the  2000  U.S.  Census  on  population  and
housing unit counts for the counties in Arizona (excluding Maricopa County and
Pima County, which are much larger than the other counties and would be placed in a
separate stratum $t = \sum_{i=1}^13 t_i$. The file has the value of $t_i$ for every county so you can calculate the population total and variance.

## a.
Calculate the selection probabilities $\psi_i$ for a sample of size 1 with probability
proportional to 2000 population. Find $\widehat{t}_{\psi}$ for each possible sample, and calculate
the theoretical variance $V(\widehat{t}_{\psi})$

## b.
Repeat (a) for an equal probability sample of size 1. How do the variances com-
pare? Why do you think one design is more efficient than the other.




We have the following formulae that we can refer to:

$$
\widehat{t}_{\psi} = \frac{1}{n}\sum_{i \in R} \frac{t_i}{\psi_i} 
$$

$$
V(\widehat{t}_\psi) = \frac{1}{n} \sum_{i=1}^N \psi_i[\frac{t_i}{\psi_i}-t]^2
$$

and

$$
\widehat{V}(\widehat{t}_\psi) =\frac{1}{n(n-1)} \sum_{i=1}^N \psi_i[\frac{t_i}{\psi_i}-\widehat{t}_\psi]^2
$$

```{r, MESSAGE=FALSE}

library(tidyverse)
library(survey)
library(srvyr)
library(pps)
library(sampling)

```

```{r, MESSAGE=FALSE}
azcounties <- read_csv('azcounties.csv')
#azcounties_clean <-azcounties %>% drop_na(population)
dim(azcounties)
#dim(azcounties_clean)
#No missing data
head(azcounties)
```
```{r}
Mi_az <- azcounties %>% group_by(name) %>% summarise(Mi = population,ti=housing)  %>% ungroup() %>% mutate(phi_eq = 1/n())

Mi_az %>% head()

Mi_az = Mi_az %>% mutate(psi_k =Mi/sum(Mi))
Mi_az
```

```{r}
#a
#is.data.frame(Mi_az)
tHat_psi <- (Mi_az$ti)/(Mi_az$psi_k) #n is 1 so we divide it by 1
t <- sum(Mi_az$ti)
I <- rep(1, 13)

tvec <-  t*I
part1<- Mi_az$psi_k*(tHat_psi-tvec)
part2<- (tHat_psi-tvec)
v.tHat <- part1 %*% part2

print("Actual total")
t

print("Estimated totals")
tHat_psi

print("Theoretical variance")
v.tHat


```

```{r}
#b
tHat_phi <- (Mi_az$ti)/(Mi_az$phi_eq) #n is one so we divide it by 1

print("Actual total")
t

print("Estimated totals - equal probability")
tHat_phi

part3 <-  Mi_az$phi_eq*(tHat_phi-tvec) # Mi_az$phi_eq= 1/13
part4 <- (tHat_phi-tvec)

v.tHat2 <- part3 %*% part4 

print("Theoretical variance - equal propbability")
v.tHat2


print("unequal prob var/equal prob var")

v.tHat/v.tHat2


```

It seems that the sample with unequal probabilities show a far better performance for both estimated mean and variance. In comparison, the sample from part(a) showed a more consistent fit for total estimates and this is reflected in the reduced variance compared to that for the smaple from part(b)  (the variance reduced to one third to that of sample with equal probabilities). This happens since samples with equal probabilities tend to overrepresent counties with small number of population such as Graham or Greenlee (a county with small population has no reason to have excessive number of housing!).


## c.
Now take a with-replacement sample of size 3. Find $\widehat{t}_{\psi}$ and $V(\widehat{t}_{\psi})$.
```{r}
set.seed(19970329)
n_az = 3  #sample we have been asked to draw

oneaz_WR <- Mi_az %>% sample_n(size=n_az, replace=T, weight=Mi)
oneaz_WR = oneaz_WR %>% group_by(name) %>% mutate(replication= 1:n())
oneaz_WR 
```
```{r, MESSAGE=FALSE}
oneaz_WR  %>% group_by(name) %>% summarise(count = n()) %>% arrange(desc(count))
oneaz_sample_WR  <-inner_join(azcounties,oneaz_WR, by='name') %>% mutate(weight= 1/(n_az*psi_k))

dim(oneaz_sample_WR)

```
```{r}
#a
#sample of size 10 with replacement
oneaz_cluster_WR_design <-svydesign(id=~name, data= oneaz_sample_WR,weight=~weight)
print("Estimated total number of housing units")
svytotal(~housing, oneaz_cluster_WR_design)
print("Actual total number of housing units")
t

print("Sample variance of estimated total")
4204.5^2

```


# 6.10

Use your sample of states drawn with probability proportional to population, from
Exercise 9, for this problem.
## a.
Using the sample, estimate the total number of counties in the United States, and
find the standard error of your estimate. How does your estimate compare with
the true value of total number of counties (which you can calculate, since the file
statepps.dat contains the data for the whole population)?
## b.
Now suppose that your friend Tom finds the ten values of numbers of counties in
your sample, but does not know that you selected these states with probabilities
proportional to population. Tom then estimates the total number of counties using
formulas for an SRS. What values for the estimated total and its standard error are
calculated by Tom? How do these values differ from yours? Is Tom’s estimator
unbiased for the population total

```{r, MESSAGE=FALSE}

library(tidyverse)
library(survey)
library(srvyr)
library(pps)
library(sampling)
```

```{r}
statepps <- read_csv('statepps.csv')

#statepps_clean <-statepps %>% drop_na(popn,counties)
#dim(statepps)
#dim(statepps_clean)
# both have dimension 51 7 -no missing data

head(statepps)
#statepps
```
```{r}
Mi_tbl <- statepps %>% group_by(state) %>% summarise(Mi = popn) %>% ungroup() %>% mutate(N = n())

Mi_tbl %>% head()
```


```{r}
Mi_tbl = Mi_tbl %>% mutate(psi_k =Mi/sum(Mi))
Mi_tbl
```

```{r}
set.seed(19970329)
n = 10  #sample we have been asked to draw

onestage_WR <- Mi_tbl %>% sample_n(size=n, replace=T, weight=Mi)
onestage_WR = onestage_WR %>% group_by(state) %>% mutate(replication= 1:n())
onestage_WR 
```
```{r}
onestage_WR  %>% group_by(state) %>% summarise(count = n()) %>% arrange(desc(count))
onestage_sample_WR  <-inner_join(statepps,onestage_WR, by='state') %>% mutate(weight= 1/(n*psi_k))

dim(onestage_sample_WR)


```
```{r}
#a
#sample of size 10 with replacement
onestage_cluster_WR_design <-svydesign(id=~state, data= onestage_sample_WR,weight=~weight)
print("Estimated total number of counties")
svytotal(~counties, onestage_cluster_WR_design)
print("Actual total number of countries")
sum(statepps$counties)
```

```{r, MESSAGE=FALSE}
#b
#SRSWR to compare
set.seed(19970319)
onestage_WR_srs <- sample_n(Mi_tbl, size=10, replace=TRUE, weight=rep(1, nrow(Mi_tbl)))
onestage_wr_sample_srs <- inner_join(statepps, onestage_WR_srs, by="state")
dim(onestage_wr_sample_srs)

```

```{r}
onestage_cluster_WR_srs_design <- svydesign(id =~state, data=onestage_wr_sample_srs, weight=~I(N/n))
svytotal(~counties, onestage_cluster_WR_srs_design)
```

#6.22

## a.

$$
E[\widehat{t}_y] = E[\sum_{i=1}^N Z_i \frac{u_i}{\pi_i}] = \sum_{i=1}^N E[Z_i \frac{u_i}{\pi_i}] = \sum_{i=1}^N E[Z_i] \frac{u_i}{\pi_i} = \sum_{i=1}^N \pi_i \frac{u_i}{\pi_i} = \sum_{i=1}^N u_i = 
$$
$$
\sum_{i=1}^N \sum_{k=1}^M \frac{l_{ik} y_k}{L_k} = \sum_{k=1}^M \sum_{i=1}^N  \frac{l_{ik}y_k}{L_k} = \sum_{k=1}^M \sum_{i=1}^N  \frac{l_{ik}y_k}{L_k} = \sum_{k=1}^M \frac{y_k}{L_k} \sum_{i=1}^N l_{ik} = \sum_{k=1}^M \frac{y_k}{L_k} L_k = \sum_{k=1}^M y_k = t_y 
$$

So we have an unbiased estimator. I.e. $E[\widehat{t}_y] = t_y$

Also,

$$
Var(\widehat{t}_y) = Var(\sum_{i=1}^N Z_i \frac{u_i}{\pi_i}) = \sum_{i=1}^N \frac{Var(Z_i)}{\pi_i^2}u_i^2 + \sum_{i=1}^N \sum_{k \neq i}^M\frac{u_i u_k}{\pi_i \pi_k} Cov(Z_i, Z_k)
$$

Since $Z_i$ is a sample indicator random variable, 

$$
Var(Z_i) = \pi_i (1-\pi_i)
$$. 

Similarly, 

$$
Cov(Z_i, Z_j) = E[Z_iZ_j] - E[Z_i]E[Z_j] = \pi_{ij} - \pi_i \pi_j
$$
$$
Var(\widehat{t}_y) =\sum_{i=1}^N \frac{\pi_i (1-\pi_i)}{\pi_i^2}u_i^2 + \sum_{i=1}^N \sum_{k \neq i}^M\frac{(\pi_{ij} - \pi_i \pi_j)}{\pi_i \pi_k} u_i u_k
$$

## b.


$$
\widehat{t}_y = \sum_{k \in S^B} \frac{1}{L^k} \sum_{i=1}^N \frac{Z_i}{\pi_i} l_{ik} y_k
$$
with weight

$$
w_k =  \frac{1}{L^k} \sum_{i=1}^N \frac{Z_i}{\pi_i} l_{ik} = \sum_{i=1}^N \frac{Z_i}{\pi_i} l_{ik} \frac{1}{L_k}
$$

while we know $\forall k \notin S^B$

$$
\sum_{i=1}^N \frac{Z_i}{\pi_i} l_{ik} y_k = 0
$$


$$
\implies w_k = \sum_{i=1}^N \frac{Z_i}{\pi_i} l_{ik} \frac{1}{L_k} = \frac{\sum_{i=1}^N \frac{Z_i}{\pi_i} l_{ik}}{\sum_{i=1}^N l_{ik}}
$$
The stuent is in $S^B$ iff the student is linoked to one of the sample units in the $S^A$.
i.e.

$$
k \in S^B \iff \sum_{i \in S^A} l_{ik} >0 
$$
$$
k \notin S^B l_{ik} = 0 \ \forall i \in S^A
$$

## c.

If $L_k =1$ we have that

$$
\widehat{t}_y = \sum_{k \in S^B} \sum_{i=1}^N \frac{l_{ik} Z_i }{\pi_i} y_k = \sum_{i=1}^N \frac{Z_i}{\pi_i} \sum_{k \in S^B} l_{ik} y_k
$$

Also, we obtain that each element $k$ belongs to exactly one unit $i$. In this case, we can view our units in the same way we have viewed PSUs.

It follows that $\sum_{k \in S^B} l_{ik}y_k$ is also the total of the values in a given PSU $i$. Therefore, it is justified to write

$$
\sum_{k \in S^B} l_{ik}y_k = t_i
$$

$$
\widehat{t}_y = \sum_{i=1}^N \frac{t_i}{\pi_i} Z_i = \widehat{t}_{HT}
$$

We have the weight of each element as

$$
w_k = \sum_{i=1}^N \frac{Z_i}{\pi_i} l_{ik} = \frac{Z_{ik}}{\pi_{ik}}
$$

Where the indicator function for unit $ik$

$$
Z_{ik} :=  \mathbf{1}_{ik}(jk) 
$$
and the inclusion probability

$$
\pi_k = P(ik \text{ is in the sample})
$$

Since unit $k$ is only associated with a unique unit $ik$, it follows that

$$
l_{jk} =0
$$

For all other $j$ in $U^A$


## d.

Remark that we are conducting SRS, we have $\pi_i = \frac{2}{3}$ $\forall \text{ unit } i$. There are three possible cases, 

$$
\begin{array}{ccc} S_1 = \{1, 2\} & S_2 = \{1,3\} & S_3 = {2, 3} \end{array}
$$
Furthermore, we have $L_k =2$ for $k = 1, 2$

Under the first case, we obtain that

$$
\widehat{t}_y = \sum_{k=1}^2 (.5) \sum_{i=1}^3 (1.5) l_{ik} y_k = .75(l_{11}+l_{12}+l_{21}+l_{22})(y_1+ y_2) = 10.5
$$

The second case:

$$
\widehat{t}_y  = .75 *(l_{21}+l_{31} + l_{22}+ l_{32}) (y_1+y_2) = 7.5
$$

and the third case:

$$
\widehat{t}_y  = .75 *(l_{11}+l_{31} + l_{12}+ l_{32}) (y_1+y_2) = 12
$$

We also remark that $\widehat{t}_y$ is unbiased.

We have

$$
\widehat{t}_y  = (.75) \sum_{k=1}^2 \sum_{i=1}^3 Z_i l_{ik} y_k
$$

$$
\implies E[\widehat{t}_y] = (.75) \sum_{k=1}^2 \sum_{i=1}^3 E[Z_i] l_{ik} y_k =\frac{3}{4}\sum_{k=1}^2 \sum_{i=1}^3 \frac{2}{3} l_{ik} y_k = \frac{1}{2}\sum_{k=1}^2 \sum_{i=1}^3  l_{ik} y_k 
$$
$$
\frac{1}{2}(l_{11}+l_{31}+ l_{21} + l_{22} + l_{12}+ l_{32})(y_1 + y_2) = (.5)(4+4+6+6) = 10
$$

From the formulae above (in part b), we are able to compute the variance
$$
\pi_{ij} = \pi_{j|i} \pi_{i} = \frac{1}{3}
$$
with $u_i$'s (denoted as a 3-size sequence)
$$
(u_i)_i = (2, 5, 3) 
$$

$$
Var(\widehat{t}_y) = 19 -15.5 =  3.5
$$


## e.

Below are the estimated values for the population total, SRSWR variance with CI's:
```{r, MESSAGE=FALSE}
library(tidyverse)
library(survey)
library(srvyr)
library(pps)
library(sampling)
library(data.table)

```
```{r}
wtshare <- read_csv('wtshare.csv')
prsc <- as.vector(wtshare$preschool)
Lk <- 1/(as.vector(wtshare$numadults)+1)
lk <- as.vector(wtshare$child)

tHat <- prsc * Lk * lk

tHaty <- 400*sum(tHat)

#Estimated total:
print("Estimated total")
tHaty
```
```{r}
setDT(wtshare)
wtshareSub <- wtshare[ , .(Total_Children = sum(preschool/(numadults+1))), by=.(Adult)]

u_i <- wtshareSub$Total_Children

SigmaSq <- ((40000*u_i)-tHaty)%*%((40000*u_i)-tHaty)/(99*100)


## Estimated Variance
print("Estimated Variance")
SigmaSq

```


```{r}
Sigma <- sqrt(SigmaSq)

L <- tHaty - (1.96*Sigma)
R <- tHaty + (1.96*Sigma)

print("95% confidence interval")
c(L, R)
```


# 6.23

## a.
Provided values suggest

$$
\begin{array}{c} \pi_1 =  \pi_{12}  + \pi_{13}  + \pi_{14} = .5\\ \pi_2 =  \pi_{21}  + \pi_{23}  + \pi_{24} = .25 \\ \pi_3 =  \pi_{31}  + \pi_{32}  + \pi_{34} = .5 \\ \pi_4 =  \pi_{41}  + \pi_{42}  + \pi_{43} = .75 \end{array}
$$

Below is the variance under w/o replacement (unequal prob)
```{r}
probs <- c(0.5, 0.25, 0.5, 0.75)
ts <- c(-5, 6, 0, -1)
beta1 <- probs*(1-probs)
beta2 <- (ts^2)/(probs^2)
vpt1 <- sum(c(beta1 * beta2))
probs2 <- c(0.004, 0.123, 0.373, 0.004, 0.123, 0.123, 0.123, 0.123, 0.254, 0.373,
0.123, 0.254)
probspairs <- c(0.5 * 0.25, 0.5 * 0.5, 0.5 * 0.75, 0.25 * 0.5,
0.25 * 0.5, 0.25 * 0.75, 0.5 * 0.5, 0.5 * 0.25, 0.5 * 0.75, 0.75 * 0.5,
0.75 * 0.25, 0.75 * 0.5)
tspairs <- c(-5 * 6, -5 * 0, -5 * -1, 6 * -5, 0, -6, 0, 0, 0, 5, -6, 0)
beta3 <- probs2-probspairs
beta4 <- tspairs/probspairs
vpt2 <- sum(c(beta3 * beta4))
VWOR <- vpt1 + vpt2
## Variance under an unequal probability sample:
VWOR

```

```{r}
psys <- (1/2) * probs
trat <- ts/psys
beta5 <- trat^2
beta6 <- psys * beta5
VWR <- (1/2) * sum(c(beta6))
## Variance under a with-replacement sample
VWR

```

SRSWR shows a lower variance, but by a very small margin; the bigger advantage comes from the fact that we may save the sampling cost.

## b.

We know that $\psi_i = \pi_i/n$ and using (6.8) it follows that

$$
n^{-1}\sum_{i=1}^N \psi_i \bigg(\frac{t_i}{\psi_i} - t \bigg)^2 = n^{-2} \sum_{i=1}^N \pi_i \bigg(\frac{nt_i}{\pi_i} - t \bigg)^2 = \sum_{i=1}^N \pi_i \bigg(\frac{nt_i}{\pi_i} - \frac{t}{n} \bigg)^2 =
$$

$$
\sum_{i=1}^N \frac{t_i^2}{\pi_i} + \sum_{i=1}^N \frac{t^2}{n^2} \pi_i - \sum_{i=1}^N \frac{t_i}{\pi_i} \frac{t}{n} \pi_i = \sum_{i=1}^N \frac{t_i^2}{\pi_i} + \frac{t^2}{n} - 2 \frac{t^2}{n} = \sum_{i=1}^N \frac{t_i^2}{\pi_i} - \frac{t^2}{n}
$$

Working from RHS:

$$
\sum_{i=1}^N\sum_{k=1}^N \pi_i \pi_k (\frac{t_i}{\pi_i} - \frac{t_k}{\pi_k})^2 = \sum_{i=1}^N\sum_{k=1}^N \pi_i \pi_k (\frac{t_i^2}{\pi_i^2} + \frac{t_k^2}{\pi_k^2} - 2\frac{t_i t_k}{\pi_i \pi_k}) = \
$$

$$
\sum_{i=1}^N\sum_{k=1}^N \frac{t_i^2}{\pi_i} \pi_k + \sum_{i=1}^N\sum_{k=1}^N \frac{t_k^2}{\pi_k} \pi_i - 2\sum_{i=1}^N t_i\sum_{k=1}^N t_k =
$$

$$
\sum_{i=1}^N \frac{t_i^2}{pi_i}\sum_{k=1}^N \pi_k + \sum_{i=1}^N \frac{t_k^2}{\pi_k} \sum_{k=1}^N \pi_i - 2t^2 = n \bigg(\sum_{i=1}^N \frac{t_i^2}{\pi_i} + \sum_{k=1}^N \frac{t_k^2}{\pi_k} \bigg) - 2t^2 = 2n\bigg( \sum_{k=1}^N  \frac{t_i^2}{\pi_i}\bigg) - 2t^2
$$
$$
\implies  \frac{1}{2n} \sum_{i=1}^N\sum_{k=1}^N \pi_i \pi_k (\frac{t_i}{\pi_i} - \frac{t_k}{\pi_k})^2 = \sum_{i=1}^N \frac{t_i^2}{\pi_i} - \frac{t^2}{n}
$$

Remarking both sides yields

$$
Var(\widehat{t}_{\psi}) = \frac{1}{2n} \sum_{i=1}^N\sum_{k=1}^N \pi_i \pi_k (\frac{t_i}{\pi_i} - \frac{t_k}{\pi_k})^2 = \frac{1}{2n} \sum_{i=1}^N\sum_{k \neq i}^N \pi_i \pi_k (\frac{t_i}{\pi_i} - \frac{t_k}{\pi_k})^2 
$$


## c.

Suppose that $\forall i, k$ $\pi_{ik} \geq \frac{n-1}{n} \pi_i \pi_k$. Remark from the previous part as well.

by (6.21) we have

$$
V(\widehat{t}_{HT}) = \frac{1}{2} \sum_{i=1}^N \sum_{k\neq i }^N (\pi_i \pi_k - \pi_{ik}) \bigg( \frac{t_i}{\pi_i} - \frac{t_k}{\pi_k} \bigg)^2 \leq \frac{1}{2} \sum_{i=1}^N \sum_{k\neq i }^N (\pi_i \pi_k - \frac{n-1}{n}\pi_{i}\pi_{k}) \bigg( \frac{t_i}{\pi_i} - \frac{t_k}{\pi_k} \bigg)^2 =
$$


$$
\frac{1}{2} \sum_{i=1}^N \sum_{k\neq i }^N (\frac{1}{n}\pi_{i}\pi_{k}) \bigg( \frac{t_i}{\pi_i} - \frac{t_k}{\pi_k} \bigg)^2 = \frac{1}{2n} \sum_{i=1}^N \sum_{k\neq i }^N (\pi_{i}\pi_{k}) \bigg( \frac{t_i}{\pi_i} - \frac{t_k}{\pi_k} \bigg)^2  =  V(\widehat{t}_\psi)
$$
In conclusion, we have shown

$$
V(\widehat{t}_{HT}) \geq  V(\widehat{t}_{\psi})
$$
## d.
We denote for a given $i$

$$
k_i = \min_{k} \frac{\pi_{ik}}{\pi_k} = \frac{\pi_{i k_{i} }}{\pi_{k_i}} = (n-1)\frac{ \pi_{i} }{n}
$$

$$
 \implies \sum_{i=1}^N \frac{\pi_{i k_i} }{ \pi_{k_i}} \geq  \frac{n-1}{n} \sum_{i=1}^N \pi_i =  n-1
$$

Meeting Gabler's condition.

## e.

$$
V(\widehat{t}_\psi) - V(\widehat{t}_{HT}) =  \frac{1}{2n} \sum_{i=1}^N\sum_{k \neq i}^N ((n-1)\pi_i \pi_k + n \pi_{ik})\bigg( \frac{t_i^2}{\pi_i^2} + \frac{t_k^2}{\pi_k^2} - 2\frac{t_i t_k}{\pi_i \pi_k} \bigg) = 
$$
and we distribute the sum.

$$
\frac{1}{2n}\sum_{i=1}^N\sum_{k \neq i}^N t_i t_k (2(n-1) - 2n \frac{\pi_{ik}}{\pi_i \pi_k}) = \sum_{i=1}^N\sum_{k \neq i}^N t_i t_k (\frac{n-1}{n} - \frac{\pi_{ik}}{\pi_i \pi_k}) \geq 0
$$

It suggests that the matrix $A = [(\frac{n-1}{n} - \frac{\pi_{ik}}{\pi_i \pi_k})_{ik}] = [a_{ik}]$ is positive semidefinite, implying its principal two by two blocks would have positive determinant.

Implying that

$$
a_{ii} a_{kk} - a_{ik}a_{ki} \geq 0 \implies (\frac{n-1}{n})^2  \geq (\frac{n-1}{n} -  \frac{\pi_{ik}}{\pi_i \pi_k})^2
$$


$$
\implies (\frac{n-1}{n})  \geq ( \frac{\pi_{ik}}{\pi_i \pi_k} - \frac{n-1}{n} ) \implies 2 (\frac{n-1}{n}) \geq \frac{\pi_{ik}}{\pi_i \pi_k} 
$$

$$
\implies  (\frac{n-1}{n} \pi_i \pi_k )\geq \pi_{ik}
$$


# 6.45 IPUMS exercise

## a. 
Select an unequal probability sampleof10 PSUs where $\psi_i \propto M_i$ where $M_i$ would be the number of persons. Take a subsample of 20 persons in each PSU


## b.
Using the sample selected, estimate the population mean and total of inctot. Give the corresponding standard errors along with the estimates. 

```{r}

library(tidyverse)
library(survey)
library(srvyr)
library(pps)
library(sampling)

```

```{r}
ipums <- read_csv('ipums.csv')
head(ipums)
dim(ipums)
#ipums_complete = ipums %>% filter(!is.na(Inctot))
#Verify that none of the data is missing
#dim(ipums_complete)
#no missing data
head(ipums)

#We look for the number of distinct PSU's
num_PSU <- ipums %>% summarize(Num_PSU=n_distinct(Psu))
num_PSU
```
```{r}
ipums_tbl <- ipums %>% group_by(Psu) %>% summarise(Mi=n()) %>% ungroup() %>% mutate(N=n())
ipums_tbl %>% head(., n=10)

```
```{r}
print("Total number of schools that will be sampled")
ipums_tbl %>% ungroup() %>% summarise(Ssutot = sum(Mi)) %>% head()
```

```{r}
set.seed(19970319)
## two-stage sampling without replacement
n=10

#Initialise

ipums_tbl = ipums_tbl %>% mutate(adj_size=Mi/sum(Mi))

#Forcing the inclusion probabilities to be \leq 1
ipums_tbl = ipums_tbl %>% mutate(pi_k = inclusionprobabilities(Mi, n=n) )

##Create vector which indicates sampled PSU's

tille_sampled <- ipums_tbl %>% with(., UPtille(pi_k))

#Filters Mi table to include only the PSU's sampled
onestage_WOR <- ipums_tbl %>% filter(tille_sampled == 1)

#Grabs SSU's from sampled PSU's

onestage_WOR_sample <- inner_join(ipums, onestage_WOR, by="Psu")

#onestage_cluster_WOR_design <-  svydesign(id=~Psu, data=onestage_WOR_sample, fpc=~pi_k, pps="brewer")
onestage_WOR_sample
```

```{r}
set.seed(19970319)
#One stage done
##Stage 2
m=20
twostage_WOR_sample <- onestage_WOR_sample %>% group_by(Psu)  %>%  sample_n(size=m, replace=F) %>% ungroup()

#twostage_WOR_sample
twostage_cluster_WOR_design <- svydesign(id=~Psu+Ssu, data=twostage_WOR_sample, weight=~I(1/pi_k)+I(Mi/m), pps="brewer")


print("Estimated population total of Inctot")
svytotal(~Inctot, twostage_cluster_WOR_design)

print("Estimated population mean of Inctot")
svymean(~Inctot, twostage_cluster_WOR_design)

```

